Effects of Reynolds Number on the Energy Conversion and Near-Wake Dynamics of a High Solidity Vertical-Axis Cross-Flow Turbine

### 2015-10-11 paper.tex
Angel --> angle; fixes #30

Hydro- or airfoil section performance is generally characterized by the
lift-to-drag
force ratio $L/D$  (or lift vs drag coefficient, $c_L/c_D$ ),
which is primarily a function of the foil's angle of attack $\alpha$. The
angles of attack
for cross-flow turbine blades are a function of turbine tip speed ratio,
induction (slowing/turning of the free stream flow prior to reaching the
turbine), and turbine rotation or azimuthal angle $\theta$. Looking at static
airfoil data, we see that the static stall angle---the angle just beyond that at
which $L/D$ is maximum---increases with blade chord Reynolds number $Re_c$
(reference?). This
is caused by the boundary layer on the foil becoming more turbulent and more
energetic, which enables it to advance further downstream against the adverse
pressure gradient on the suction side and delay separation to higher angels of
attack, as well as diminish the negative effects of a laminar separation
bubble occurring near the leading edge (reference?). For smooth airfoils, this
transition can
cause a dramatic increase in foil performance, as seen in
Figure~\ref{fig:McMasters} at a blade chord Reynolds number on the order of
$10^5$ \cite{McMasters1980}. A review of low Reynolds number airfoil behavior is
presented in \cite{Lissaman1983}. Note that there is a distinct lack of highly
reliable data for airfoils in this transitional regime and below. An evaluation
of the various databases relevant to cross-flow turbines is presented in
\cite{Bedon2014}.

>>>>

Hydro- or airfoil section performance is generally characterized by the
lift-to-drag
force ratio $L/D$  (or lift vs drag coefficient, $c_L/c_D$ ),
which is primarily a function of the foil's angle of attack $\alpha$. The
angles of attack
for cross-flow turbine blades are a function of turbine tip speed ratio,
induction (slowing/turning of the free stream flow prior to reaching the
turbine), and turbine rotation or azimuthal angle $\theta$. Looking at static
airfoil data, we see that the static stall angle---the angle just beyond that at
which $L/D$ is maximum---increases with blade chord Reynolds number $Re_c$
(reference?). This
is caused by the boundary layer on the foil becoming more turbulent and more
energetic, which enables it to advance further downstream against the adverse
pressure gradient on the suction side and delay separation to higher angles of
attack, as well as diminish the negative effects of a laminar separation
bubble occurring near the leading edge (reference?). For smooth airfoils, this
transition can
cause a dramatic increase in foil performance, as seen in
Figure~\ref{fig:McMasters} at a blade chord Reynolds number on the order of
$10^5$ \cite{McMasters1980}. A review of low Reynolds number airfoil behavior is
presented in \cite{Lissaman1983}. Note that there is a distinct lack of highly
reliable data for airfoils in this transitional regime and below. An evaluation
of the various databases relevant to cross-flow turbines is presented in
\cite{Bedon2014}.
